Comparison of dimensionality reduction schemes for parallel global optimization algorithms

### 2018-05-10 2018/YSC/ysc_paper.tex
YSC: refactoring in bibliography

In the present paper, the parallel algorithms for solving the multiextremal optimization problems
are considered. In the multiextremal problems, the opportunity of reliable estimate of the global
optimum is based principally on the availability of some information on the function known
{\textit a priori} allowing relating the probable values of the optimized function to the known
values at the points of performed trials. Very often, such an information on the problem being
solved is represented in the form of suggestion that the objective function $\varphi(y)$ satisfies
Lipschitz condition with the constant $L$ not known a priori (see, for example,
\cite{Jones,Gablonsky,Evtushenko}). At that, the objective function could be defined by a
program code i. e. could represent a ``black-box''-function. Such problems are presented in the
applications widely (problems of optimal design of objects and technological processes in
various fields of technology, problems of model fitting according to observed data in scientific
research, etc.).

@@

In Sec.~\ref{sec:results_parallel} the results of the numerical experiments are presented, which
have demonstrated the algorithm from Sec.~\ref{sec:parallel_evolvents} based on the rotated
evolvents allowed obtaining the speedup up to 43 times when solving the problem series
employing several nodes of the computer cluster. It is worth noting that the objective functions
in the considered problems are not computation-costly (the averaged computation time was
$10^{-3}$ sec). In the case of more complex problems,the speedup in time could approach the
speedup with respect to the number of iterations.
%------------------------------------------------------------------------------
% Index
%\printindex
\begin{thebibliography}{107}

  \bibitem{sergeyevStronginLera2013}% (MR3113120) [10.1007/978-1-4614-8042-6]
  \newblock Y. D. Sergeyev, R. G. Strongin and D. Lera,
  \newblock \emph{Introduction to Global Optimization Exploiting Space-filling Curves},
  \newblock Springer (2013)

  \bibitem{strongin1978}% (MR509033)
  \newblock R. G. Strongin,
  \newblock \emph{Numerical Methods in Multi-Extremal Problems (Information-Statistical Algorithms)},
  \newblock Moscow: Nauka (1978) (In Russian)

  \bibitem{Strongin1992}% (MR1263606) [10.1007/BF00122428]
  \newblock R. G. Strongin,
  \newblock \emph{\emph{Algorithms for multi-extremal mathematical programming problems
  employing a set of joint space-filling curves}},
  \newblock \emph{J. Glob. Optim.}, \textbf{2}, 357--378 (1992)

  \bibitem{stronginGergelBarkalovParGO}
  \newblock R. G. Strongin, V. P. Gergel, V. A. Grishagin and K. A. Barkalov,
  \newblock \emph{Parallel Computations for Global Optimization Problems},
  \newblock Moscow State University, Moscow (2013) (In Russian)

  \bibitem{strSergGO}%5 (MR1797058) [10.1007/978-1-4615-4677-1]
  \newblock R. G. Strongin and Y. D. Sergeyev,
  \newblock \emph{Global Optimization with Non-convex Constraints. Sequential and Parallel
  Algorithms},
  \newblock Kluwer Academic Publishers, Dordrecht (2000, 2nd ed. 2013, 3rd ed. 2014)

  \bibitem{zilinskTornGO}% (MR1100586) [10.1007/3-540-50871-6]
  \newblock A. T\"orn and A. \v Zilinskas,
  \newblock \emph{Global Optimization},
  \newblock Springer, (1989)

  \bibitem{Goryachih2017}
  \newblock Goryachih, A.
  \newblock \emph{A class of smooth modification of space-filling curves for global optimization
  problems}
  \newblock Springer Proceedings in Mathematics and Statistics, \textbf{197}, pp. 57--65 (2017)

  \bibitem{zhigljavskyRandGO}% (MR1187048) [10.1007/978-94-011-3436-1]
  \newblock A. A. Zhigljavsky,
  \newblock \emph{Theory of Global Random Search},
  \newblock Kluwer Academic Publishers, Dordrecht (1991)

  \bibitem{Strongin1991}
  Strongin, R.G.: \emph{Parallel multi-extremal optimization using a set of evolvents}. Comp. Math.
  Math. Phys. \textbf{31(8)}, 37--46 (1991)

  \bibitem{Gergel2009}
  Strongin, R.G., Gergel, V.P., Barkalov, K.A.: \emph{Parallel methods for global optimization problem
  solving.} Journal of instrument engineering. \textbf{52}, 25--33 (2009) (In Russian)

  \bibitem{Gaviano2003}
  Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: \emph{Software for generation of classes
  of test functions with known local and global minima for global optimization.} ACM
  Transactions on Mathematical Software \textbf{29(4)}, 469--480 (2003)

  \bibitem{grishaginClass}
  Grishagin, V.A.: \emph{Operating Characteristics of Some Global Search Algorithms.} Problems of
  Statistical Optimization \textbf{7}, 198--206 (1978) (In Russian)

  \bibitem{globalizerSystem}
  Gergel V.P., Barkalov K.A., and Sysoyev A.V: Globalizer: \emph{A novel supercomputer software
  system for solving time-consuming global optimization problems.} Numerical Algebra, Control
  \& Optimization \textbf{8(1)}, 47--62 (2018)

  \bibitem{Sergeyev2006}
  Sergeyev, Ya.D., Kvasov, D.E.: \emph{Global search based on efficient diagonal partitions and a set
  of Lipschitz constants}. SIAM J. Optim \textbf{16(3)}, 910--937 (2006)

  \bibitem{SergeyevKvasov2015}
  Sergeyev, Y.D., Kvasov, D.E.: \emph{A deterministic global optimization using smooth diagonal
  auxiliary functions.} Communications in Nonlinear Science and Numerical Simulation. \textbf{21(1-3)},
  99--111 (2015)

  \bibitem{Zilinskas2008}
  \v Zilinskas, J.: \emph{Branch and bound with simplicial partitions for global optimization.} Math.
  Model. Anal. \textbf{13(1)}, 145--159 (2008)

  \bibitem{Zilinskas2014}
  Paulavi\v cius, R., \v Zilinskas, J.: \emph{Simplicial Lipschitz optimization without the Lipschitz
  constant.} J. Glob. Optim. \textbf{59(1)}, 23-40 (2014)

  \bibitem{Jones}
  Jones, D.R.: The direct global optimization algorithm. In: Floudas, C.A., Pardalos, P.M.
  (eds.) \emph{The Encyclopedia of Optimization}, 2nd edn., pp. 725--735. Springer, Heidelberg
  (2009)

  \bibitem{Gablonsky}
  Gablonsky, J.M., Kelley, C.T.: \emph{A locally-biased form of the DIRECT algorithm.} J. Glob.
  Optim., \textbf{21(1)}, 27--37 (2001)

  \bibitem{Evtushenko}
  Evtushenko, Y., Posypkin, M.: \emph{A deterministic approach to global box-constrained
  optimization.} Optim. Lett. \textbf{7(4)}, 819--829 (2013)
\end{thebibliography}

>>>>

In the present paper, the parallel algorithms for solving the multiextremal optimization problems
are considered. In the multiextremal problems, the opportunity of reliable estimate of the global
optimum is based principally on the availability of some information on the function known
{\textit a priori} allowing relating the probable values of the optimized function to the known
values at the points of performed trials. Very often, such an information on the problem being
solved is represented in the form of suggestion that the objective function $\varphi(y)$ satisfies
Lipschitz condition with the constant $L$ not known a priori (see, for example,
\cite{Jones2009,Gablonsky2001,Evtushenko2013}). At that, the objective function could be defined by a
program code i. e. could represent a ``black-box''-function. Such problems are presented in the
applications widely (problems of optimal design of objects and technological processes in
various fields of technology, problems of model fitting according to observed data in scientific
research, etc.).

@@

In Sec.~\ref{sec:results_parallel} the results of the numerical experiments are presented, which
have demonstrated the algorithm from Sec.~\ref{sec:parallel_evolvents} based on the rotated
evolvents allowed obtaining the speedup up to 43 times when solving the problem series
employing several nodes of the computer cluster. It is worth noting that the objective functions
in the considered problems are not computation-costly (the averaged computation time was
$10^{-3}$ sec). In the case of more complex problems,the speedup in time could approach the
speedup with respect to the number of iterations.
%------------------------------------------------------------------------------
% Index
%\printindex
%\label{sect:bib}
\bibliographystyle{unsrt}
\bibliography{ysc_paper}
